---
title: "p8105_hw3_rl3401"
author: "Ruoxi Li"
date: "`r Sys.Date()`"
output: github_document
---
```{r}
library(tidyverse)
```



##problem 2

```{r}
library(p8105.datasets)
data("brfss_smart2010") 
```

```{r}
brfss_df <- brfss_smart2010 |>
  janitor::clean_names() |>
  filter( topic== "Overall Health",
          response %in% c("Excellent","Very good","Good","Fair","Poor")
            ) |>
  mutate(response = fct_relevel(response, "Poor", "Fair", "Good", "Very good", "Excellent"))
```


```{r}
brfss_df |>
  group_by(year,locationabbr)|>
  summarize(n_obs = n()) |>
  filter(year == 2002, n_obs >= 7)
```

In 2002, which states were observed at 7 or more locations? 36.

```{r}
brfss_df |>
  group_by(year,locationabbr)|>
  summarize(n_obs = n()) |>
  filter(year == 2010, n_obs >= 7)
```

What about in 2010? 45.

Construct a dataset that is limited to Excellent responses, and contains, year, state, and a variable that averages the data_value across locations within a state. Make a “spaghetti” plot of this average value over time within a state (that is, make a plot showing a line for each state across years – the geom_line geometry and group aesthetic will help).

```{r}
excellent_df = 
  brfss_df|>
  filter(response == "Excellent") 
excellent_df |>
  group_by(year,locationabbr)|>
  summarize(mean_data_value = mean(data_value, na.rm= TRUE)) |>    
  ggplot(aes(x=year,y=mean_data_value,group=locationabbr))+
  geom_line()+
  labs(x = "Year", y = "Average of the data value", title = "“spaghetti” plot of this average value over time within a state ")
```


Make a two-panel plot showing, for the years 2006, and 2010, distribution of data_value for responses (“Poor” to “Excellent”) among locations in NY state

```{r}
brfss_df |>
  filter(locationabbr == "NY", year %in% c(2006,2010))|>
  ggplot(aes(x = data_value, color = response)) +
  facet_grid(.~ year) +
  geom_density(alpha = 0.5) +
  labs(x = "Data Value", y = "Count", title = "Distribution of Data Value by Year in NY State") 
```

## problem 3

```{r}
accel_df = read_csv("data/nhanes_accel.csv")|>
  janitor::clean_names()

covar_df = read_csv("data/nhanes_covar.csv",skip = 4)|>
  janitor::clean_names()
```

```{r}
baseline_df =
covar_df |>
  drop_na()|>
  filter(age >= 21)|>
  mutate(
      sex = case_match(
      sex,
      1 ~ "male",
      2 ~ "female"),
      education = case_match(
        education,
        1 ~ "less than high school",
        2 ~ "high school equivalent",
        3 ~ "more than high school"
      ))|>
  mutate(education = fct_relevel(education, "less than high school", "high school equivalent", "more than high school"))
```
Produce a reader-friendly table for the number of men and women in each education category
```{r}
baseline_df |>
  group_by(sex,education)|>
  summarize(n_obs=n())|>
  pivot_wider(
    names_from = education,
    values_from = n_obs
  )|>
  knitr::kable()
```

and create a visualization of the age distributions for men and women in each education category. Comment on these items.

```{r}
baseline_df|>
  ggplot(aes(x=age,color=sex))+
  geom_density()+
  facet_grid(.~education)
```

```{r}
total_activity_df = accel_df|>
  pivot_longer(
    cols = starts_with("min"),
               names_to = "time",
               values_to = "accel")|>
  group_by(seqn)|>
  summarize(total_activity=sum(accel))
```

```{r}
total_activity_plot_df = inner_join(baseline_df,total_activity_df,by="seqn") 
total_activity_plot_df|>
  ggplot(aes(x=age,y=total_activity,color=sex))+
  facet_grid(.~education)+
  geom_point()+
  geom_smooth()
```


```{r}
hour_df = accel_df|>
   pivot_longer(
    cols = starts_with("min"),
               names_to = "time",
               values_to = "accel")|>
   mutate(hour = case_when(
    time %in% paste0("min", 1:60) ~ "hour1",
    time %in% paste0("min", 61:120) ~ "hour2",
    time %in% paste0("min", 121:180) ~ "hour3",
    time %in% paste0("min", 181:240) ~ "hour4",
    time %in% paste0("min", 241:300) ~ "hour5",
    time %in% paste0("min", 301:360) ~ "hour6",
    time %in% paste0("min", 361:420) ~ "hour7",
    time %in% paste0("min", 421:480) ~ "hour8",
    time %in% paste0("min", 481:540) ~ "hour9",
    time %in% paste0("min", 541:600) ~ "hour10",
    time %in% paste0("min", 601:660) ~ "hour11",
    time %in% paste0("min", 661:720) ~ "hour12",
    time %in% paste0("min", 721:780) ~ "hour13",
    time %in% paste0("min", 781:840) ~ "hour14",
    time %in% paste0("min", 841:900) ~ "hour15",
    time %in% paste0("min", 901:960) ~ "hour16",
    time %in% paste0("min", 961:1020) ~ "hour17",
    time %in% paste0("min", 1021:1080) ~ "hour18",
    time %in% paste0("min", 1081:1140) ~ "hour19",
    time %in% paste0("min", 1141:1200) ~ "hour20",
    time %in% paste0("min", 1201:1260) ~ "hour21",
    time %in% paste0("min", 1261:1320) ~ "hour22",
    time %in% paste0("min", 1321:1380) ~ "hour23",
    time %in% paste0("min", 1381:1440) ~ "hour24",
    TRUE ~ as.character(time)
  )
   ) |>
    group_by(seqn,hour)|>
  summarize(hour_total=sum(accel))

hour_plot_df = inner_join(baseline_df, hour_df, by = "seqn")


hour_plot_df |>
ggplot( aes(x = hour, y = hour_total, color = sex)) +
  geom_line(aes(group = seqn), alpha = 0.5)   +
  geom_smooth() +
  facet_grid(.~education)+
  labs(title = "24-hour activity time courses by education level",
       x = "Hour",
       y = "Activity Level",
       color = "Sex") 
```
