---
title: "p8105_hw3_rl3401"
author: "Ruoxi Li"
date: "`r Sys.Date()`"
output: github_document
---
```{r}
library(tidyverse)
```

##problem 2

```{r}
library(p8105.datasets)
data("brfss_smart2010") 
```

```{r}
brfss_df <- brfss_smart2010 |>
  janitor::clean_names() |>
  filter( topic== "Overall Health",
          response %in% c("Excellent","Very good","Good","Fair","Poor")
            ) |>
  mutate(response = fct_relevel(response, "Poor", "Fair", "Good", "Very good", "Excellent"))
```


```{r}
brfss_df |>
  group_by(year,locationabbr)|>
  summarize(n_obs = n()) |>
  filter(year == 2002, n_obs >= 7)
```

In 2002, which states were observed at 7 or more locations? 36.

```{r}
brfss_df |>
  group_by(year,locationabbr)|>
  summarize(n_obs = n()) |>
  filter(year == 2010, n_obs >= 7)
```

What about in 2010? 45.

Construct a dataset that is limited to Excellent responses, and contains, year, state, and a variable that averages the data_value across locations within a state. Make a “spaghetti” plot of this average value over time within a state (that is, make a plot showing a line for each state across years – the geom_line geometry and group aesthetic will help).

```{r}
excellent_df = 
  brfss_df|>
  filter(response == "Excellent") 
excellent_df |>
  group_by(year,locationabbr)|>
  summarize(mean_data_value = mean(data_value, na.rm= TRUE)) |>    
  ggplot(aes(x=year,y=mean_data_value,group=locationabbr))+
  geom_line()+
  labs(x = "Year", y = "Average of the data value", title = "“spaghetti” plot of this average value over time within a state ")
```


Make a two-panel plot showing, for the years 2006, and 2010, distribution of data_value for responses (“Poor” to “Excellent”) among locations in NY state

```{r}
brfss_df |>
  filter(locationabbr == "NY", year %in% c(2006,2010))|>
  ggplot(aes(x = data_value, color = response)) +
  facet_grid(.~ year) +
  geom_density(alpha = 0.5) +
  labs(x = "Data Value", y = "Count", title = "Distribution of Data Value by Year in NY State") 
```

## problem 3


